Sketch Recogniton using Particle Swarm Algorithm

### 2009-04-12 draft/System.tex
after correctly splled now look at any thing else

 In this research, time difference, direction, speed and curvature of each point along the stroke was computed. The experiments in section \ref{} proved that computing all information's provide better segmentations results than only using one of the graphs. The time difference is calculated as $\Delta t = t_{i+1} - t_i$ where $t_i$ is the time of the point $i$ and $t_{i+1}$ is the time of the point $i+1$. The speed is calcuated using equation $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.

@@

\subsection{Critical Point Detection}
\label{sec:CriticalPointDetection}
After computing speed, time difference and curvature information the system proceeds to detect the points with low velocity and high curvatures. Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated. Then a threshold \textit{th} is used to separate the curve into regions; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and the dominate points $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.  % (as shown are redundant)
\begin{figure}
	\centering
		\includegraphics[scale=0.5]{images/speedreg.eps}
	\caption{Possible Dominant Points Calculations}
	\label{fig:MaxRegioi}
\end{figure}

@@

The system repeats this process to curvature, direction, time difference and speed curves. For each curve the points labeled as possible dominant points $P_{pd}$ are saved into a single array. The points that are repeatedly labeled are given a higher score than the other points. This score is then used to sort the  $P_{pd}$ points array. The array that contains possible points $P_{pd}$ is used as input for the next segmentation step. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominate points $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.

\begin{figure}
	\centering
		\includegraphics{images/ppd.eps}
	\caption{Posible domninate points}% a) Possible dominate point b) Particle encoding  
	\label{fig:LabelsPPD}
\end{figure}

@@

The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set of dominate points that defines the stroke with minimum segmentation error. In other words, the system tries to find the fewer number of points (1's) location in the particle array which gives minimum segmentation error.

@@

\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
\label{sec:PolygonDivisionAlgorithm}
The first algorithm, the algorithm tries to segment the stroke as a polygon. The final solution is a set of line segments that best define the input stroke. Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end. The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
 where $M$ is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}. Figure \ref{fig:DPSOERROR} shows a graphical representation of the error. The fitness is computed using equation \ref{eq:fitnessSwarm1} 
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.  If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.  This fitness function optimize two goals the first is to move the solution into the feasible solution space with acceptable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices'\cite{PolygonApproximationPSO}. 
\begin{figure}
	\centering
  \includegraphics[scale=0.8]{images/pso1.eps}			
	\caption{ALS1 Error}% a) Possible dominate point b) Particle encoding  
	\label{fig:DPSOERROR}
\end{figure}
 %The error is chosen to in favor of larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices'.\\ %we will want to lower the number of vertices'. \\
%if we say that s
%Alg1:
%Alg2: 
\subsubsection{Curve Divisions Algorithm \textsl{AlgS2}}
\label{sec:PolygonDivisionAlgorithm}

@@

\paragraph{Determinig The Segment Type.} 
\label{sec:DeterminigSegmentType}
If the average distance $d_s$ equals zero this means that the segment$\widehat{x_ix_j}$ points is exactly the straight line approximated. If the distance ($d_c$) equals zero means that the segments point lies exactly on the circular arc estimated. But the user rarely draws an exact line or circle, so the distances are used as the fitting error. The minimum distance determines the type of the current segment, in other words if ($d_s$) is smaller than ($d_c$) then the segment is labeled as a line segment otherwise it is labeled as a circular arc.

@@

\subsubsection{Solution Refine procedures } 
%\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
 After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution. Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{sec:Preprocessing}), if not the point is moved to the nearest labeled point.
   After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure. For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{sec:Preprocessing}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominate points $P_{pd}$. After that the particles are tested to make sure that the distance between every two successive dominate point is larger than the constant $min_D$. If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.

\subsection{Choosing the best fit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error($E_{alg}$ in Equation \ref{eq:bestFit2})represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to  compare the two segmentation fairly. 
\begin{equation}
\label{eq:bestFit1}
E_{area}=A_{st}-A_{sg}
 \end{equation}
 \begin{equation}
\label{eq:bestFit2}
E_{tot}=E_{area}+E_{alg}
 \end{equation}
 where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmenation, $E_{alg}$ is the Segmentation Error. The segmentation which crosspond to the minimum $E_{tot}$ is the best segmentation.

\section{Hybird Segmentation Algorithm \textsl{Alg3}}
\label{sec:BenchMarckAlgorithm}

@@

\begin{description}
 \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
\begin{description}
	 \item[Segments] Number of segments in the symbol.
	 \item [Strokes] Number of strokes or partical strokes that created the symbols.  
		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item [Curves] Number of curves or ellipses in the symbol. 
		\item [Lines] Number of lines in the symbol. 
		\item [Perpendicular lines] Number of perpendicular lines. 
		\item [Parallel lines] Number of parallel lines. 
		\item [Intersections] Number of intersection between lines and curves. 
		\item [T intersections] Number of T intersections. 
		\item [L intersections] Number of L intersections. 
		\item [X intersections] Number of X intersections.
		\item [Size Ratio] Ratio between width to height of the symbol.

@@

\section{Classification}%%Feature Set
\label{sec:Classification}
 The system uses support vector machine (SVM) classifier with Gaussian kernel (RBF kernel) \cite{libsvm}. The system uses OVO classifiers structure (object versus object). A validation algorithm is used to choose the Kernel variables parameters $c, \gamma$. The Rubine algorithm (Linear Gaussian classifier) was implemented for comparing the system. The results Section \ref{sec:recognitionAlgorithms} demonstrates the results of both classifiers. \\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points.

>>>>

 In this research, time difference, direction, speed and curvature of each point along the stroke was computed. The experiments in section \ref{sec:PSO} proved that computing all information's provide better segmentations results than only using one of the graphs. The time difference is calculated as $\Delta t = t_{i+1} - t_i$ where $t_i$ is the time of the point $i$ and $t_{i+1}$ is the time of the point $i+1$. The speed is calculated using equation $v=\Delta s/\Delta t$ where $t$ is the time difference between two points and $s$ is the length between them.

@@

\subsection{Critical Point Detection}
\label{sec:CriticalPointDetection}
After computing speed, time difference and curvature information the system proceeds to detect the points with low velocity and high curvatures. Using simple differentiation to detect local extreme points resulted in false points due to the non smooth curves. Hence, the system adopted a process presented by \cite{earlyprocess}, where the mean of the curve is calculated. Then a threshold \textit{th} is used to separate the curve into regions; each region $Region_i$ is defined as a range of points, where the curve values are either above or below the threshold \textit{th}. Those regions are further processed to find the maximum point $Max(Region_i)$ of each region $Region_i$. The stroke points $p_i(x,y)$ that correspond to those maximum values are labeled as \textit{possible dominant points} $P_{pd}$. Figure \ref{fig:MaxRegioi} shows an example of $Region_i$ of the speed curve and thedominantpoints $P_{pd}$ which correspond to the minimum of point in this region $Min(Region_i)$.  % (as shown are redundant)
\begin{figure}
	\centering
		\includegraphics[scale=0.5]{images/speedreg.eps}
	\caption{Possible Dominant Points Calculations}
	\label{fig:MaxRegioi}
\end{figure}

@@

The system repeats this process to curvature, direction, time difference and speed curves. For each curve the points labeled as possible dominant points $P_{pd}$ are saved into a single array. The points that are repeatedly labeled are given a higher score than the other points. This score is then used to sort the  $P_{pd}$ points array. The array that contains possible points $P_{pd}$ is used as input for the next segmentation step. Figure \ref{fig:LabelsPPD} shows the particles labeled as Possible dominant points $P_{pd}$ by the preprocessing, it is noticed the redundancy of some $P_{pd}$ points.

\begin{figure}
	\centering
		\includegraphics{images/ppd.eps}
	\caption{Possible dominant points}% a) Possibledominantpoint b) Particle encoding  
	\label{fig:LabelsPPD}
\end{figure}

@@

The goal of the DPSO algorithm is to find the solution $P_i$ that generates the minimum set ofdominantpoints that defines the stroke with minimum segmentation error. In other words, the system tries to find the fewer number of points (1's) location in the particle array which gives minimum segmentation error.

@@

\subsubsection{Polygon Division Algorithm \textsl{AlgS1}}
\label{sec:PolygonDivisionAlgorithm}
The first algorithm, the algorithm tries to segment the stroke as a polygon. The final solution is a set of line segments that best define the input stroke. Given the inputs stroke let's define the set of points in the stroke as  $S = \left\{ {x_1 ,x_2  \ldots x_N } \right\}$ which is the set of consecutive points from start of stroke till the end. The arc $\widehat{x_ix_j}$ is defined as the consecutive set of points from point $x_i,x_{i+1} \cdots,x_j$. The line
$\overline{x_i x_j} $ is the straight line connecting point $x_i$ to point $x_j$. The approximation error is computed by the equation \ref{eq:ErrorSwarm1} 
\begin{equation}
E=\sum\nolimits_{i = 0}^M e ( \widehat{x_ix_{i+1}},\overline{x_i x_{i+1}})
\label{eq:ErrorSwarm1}
\end{equation}
 where $M$ is the number of dominant points in this solution.  The error $ e ( \widehat{x_ix_j},\overline{x_i x_j})$ is computed as the sum of squared perpendicular distance from every point along the arc $\widehat{x_ix_j}$ to the line $\overline{x_i x_j}$ \cite{PolygonApproximationPSO}. Figure \ref{fig:DPSOERROR} shows a graphical representation of the error. The fitness is computed using equation \ref{eq:fitnessSwarm1} 
\begin{equation}
\max fitness(p_i ) = \left\{ {\begin{array}{*{20}c}
   { - E/\varepsilon N} & {ifE > \varepsilon ,}  \\
   {D/\sum\limits_{j = 1}^N {p_{ij} } } & {otherwise}  \\
\end{array}} \right.
\label{eq:fitnessSwarm1}
\end{equation}%\]
where $N$ is the number of points in the stroke, $D$ is the number of point in the solution that was previously labeled as $P_{pd}$  and $E$ is the computed error and $\varepsilon$ is the error threshold.  If the solution produced an error that exceed the error threshold $\varepsilon$ the fitness function assign a -ve value for the solution to express it's in feasibility, otherwise the inverse of the number of vertex produced is used to access the solution fitness in terms of minimum number of vertices'.  This fitness function optimize two goals the first is to move the solution into the feasible solution space with acceptable error bounds and second is to fly the particle to a new position which may result in a polygon with a fewer number of vertices'\cite{PolygonApproximationPSO}. 
\begin{figure}
	\centering
  \includegraphics[scale=0.8]{images/pso1.eps}			
	\caption{AlgS1 Error}% a) Possibledominantpoint b) Particle encoding  
	\label{fig:DPSOERROR}
\end{figure}
 %The error is chosen to in favor of larger than the threshold is given a -ve value to lower the value of solution otherwise the system will favor the lower number of vertices'.\\ %we will want to lower the number of vertices'. \\
%if we say that s
%Alg1:
%Alg2: 
\subsubsection{Curve Divisions Algorithm \textsl{AlgS2}}
\label{sec:PolygonDivisionAlgorithm}

@@

\paragraph{Determining The Segment Type.} 
\label{sec:DeterminigSegmentType}
If the average distance $d_s$ equals zero this means that the segment$\widehat{x_ix_j}$ points is exactly the straight line approximated. If the distance ($d_c$) equals zero means that the segments point lies exactly on the circular arc estimated. But the user rarely draws an exact line or circle, so the distances are used as the fitting error. The minimum distance determines the type of the current segment, in other words if ($d_s$) is smaller than ($d_c$) then the segment is labeled as a line segment otherwise it is labeled as a circular arc.

@@

\subsubsection{Solution Refine procedures } 
%\cite{PolygonApproximationPSO} used a merge and divide algorithm after each loop of the swarm system to refine the solution but we used another enhancement method.
 After each loop in the swarm algorithm, each particle loops on the set of selected dominant points to enhance the solution. Each dominant point is checked to find if it was labeled before as a possible dominant (computed as in section \ref{sec:Preprocessing}), if not the point is moved to the nearest labeled point.
   After each loop of the swarm algorithm (\textsl{AlgS1} and \textsl{AlgS2}), each particle is refined using the following procedure. For each particle $P_i$ each dominant point $P_{ij}$ is checked to find if it was labeled before as a \textit{possible dominant point} $P_{pd}$ (computed as in section \ref{sec:Preprocessing}). If it was not labeled the point $P_{ij}$ is moved to the nearest labeled point. This ensures that all of the points generated by the DPSO are possible dominant points $P_{pd}$. After that the particles are tested to make sure that the distance between every two successive dominant point is larger than the constant $min_D$. If two points are nearer than $min_D$\footnote{Constant selected to be 10 pixels} then one of the points is removed.

\subsection{Choosing the best fit}
The two PSO algorithms produce two segmentation solutions. After that the system evaluates each solution to finally select the best segmentation. This evaluation is based on the area of both the segmentation and the input stroke. The segmentation error($E_{alg}$ in Equation \ref{eq:bestFit2})represents the sum of Square Root Error of the length from the point on the stroke to the estimated curve or line. In \textsl{AlgS2}, this error is the same as the computed error in Equation \ref{eq:ErrorSwarm1} but For the second algorithm \textsl{AlgS2} it is different than the computed error in Equation \ref{eq:errorSwarm2}. The additional computation is necessary to  compare the two segmentation fairly. 
\begin{equation}
\label{eq:bestFit1}
E_{area}=A_{st}-A_{sg}
 \end{equation}
 \begin{equation}
\label{eq:bestFit2}
E_{tot}=E_{area}+E_{alg}
 \end{equation}
 where $A_{st}$ is the area of the input stroke, $A_{sg}$ is area of computed segmentation, $E_{alg}$ is the Segmentation Error. The segmentation which cross ponds to the minimum $E_{tot}$ is the best segmentation.

\section{Hybrid Segmentation Algorithm \textsl{Alg3}}
\label{sec:BenchMarckAlgorithm}

@@

\begin{description}
 \item [Structural and geometrical Features(FS1)] Features defines the structure of the geometrical symbol.  
\begin{description}
	 \item[Segments] Number of segments in the symbol.
	 \item [Strokes] Number of strokes or partial strokes that created the symbols.  
		\item[Primitives] Number of primitives in the symbol. The feature helps when identifying             symbols with mixed geometric primitives like cylinders and callouts.  
		\item [Curves] Number of curves or ellipses in the symbol. 
		\item [Lines] Number of lines in the symbol. 
		\item [Perpendicular lines] Number of perpendicular lines. 
		\item [Parallel lines] Number of parallel lines. 
		\item [Intersections] Number of intersection between lines and curves. 
		\item [T intersections] Number of T intersections. 
		\item [L intersections] Number of L intersections. 
		\item [X intersections] Number of X intersections.
		\item [Size Ratio] Ratio between width to height of the symbol.

@@

\section{Classification}%%Feature Set
\label{sec:Classification}
 The system uses support vector machine (SVM) classifier with Linear kernel (Linear kernel) \cite{libsvm}. The system uses OVO classifiers structure (object versus object). A validation algorithm is used to choose the Kernel variables parameters $c$ The results Section \ref{sec:recognitionAlgorithms} demonstrates the results of both classifiers. \\%Some features like number of parallel, perpendicular and intersecting lines are computed from the segmentation other like density, centroid and Zernike moments are computed from the raw stroke points. . The Rubine algorithm (Linear Gaussian classifier) was implemented for comparing the system.
